Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees

### 2023-02-20 script/manuscript_MEE.Rmd
catching weirds

  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Solomon'
duplicate_reps_in_sample <- duplicate_reps_in_sample %>% 
  ungroup() %>% 
  group_by(sample.id, Genus) %>% 
  mutate(Prcnt_seqs_Reclass = 
           ifelse(n() == 2, Prcnt_seqs/2, sum(Prcnt_seqs)/n_distinct(plant.species)/2 ),
         TAXON_EXPERT = plant.species) %>% 
  distinct(sample.id, TAXON_EXPERT, .keep_all = T)

@@

  mutate(TAXON_EXPERT = case_when(
    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 4 ~ plant.species, # 4 has a viola split

@@

    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', 
    sample.id == 19 ~ plant.species,

    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
    sample.id == 20 ~ plant.species,
    sample.id == 21 ~ plant.species,

    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 23 ~ plant.species, # has both Viola

    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,

    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,

    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,

    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 27 ~ plant.species, # has both Hydrophyllum

    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,

    sample.id == 29 ~ plant.species,

    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 30 ~ plant.species,

    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 32 ~ plant.species,

    sample.id == 33 ~ plant.species, # has both Viola

@@

    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 39 ~ plant.species, # has both hydrophyllum
    sample.id == 40 ~ plant.species,

    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
    sample.id == 41 ~ plant.species,

    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,
    sample.id == 44 ~ plant.species, # 44 has both delphinium

@@

    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 51 ~ plant.species, # 51 has a viola split

    sample.id == 52 ~ plant.species,
    sample.id == 54 ~ plant.species,

    ######### 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? ####
    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 55 ~ plant.species,

@@

   TRUE ~ as.character(TAXON_NEW)
  )) %>% 
  mutate(
    TAXON_EXPERT = str_replace_all(TAXON_EXPERT, '[.]', ' '),
    TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Arenaria globiflora', 'Eremogone congesta'),
    TAXON_EXPERT = if_else(doy < 190 & TAXON_EXPERT == 'Lupinus argenteus', 
                           'Lupinus sericeus', TAXON_EXPERT),) %>%

@@

# now we split up the reads assigned to conflicted groups  # e.g. Cynoglossum amplifolium in sample 65
b <- blst_reclass_expert %>% 
  filter(sample.id %in% dupe_samples) %>% 
  group_by(sample.id, Family) %>% 
  filter(n() >= 3)

>>>>

  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Solomon'
duplicate_reps_in_sample <- duplicate_reps_in_sample %>% 
  ungroup() %>% 
  group_by(sample.id, Genus) %>% 
  mutate(Prcnt_seqs_Reclass = 
           ifelse(n() == 2, Prcnt_seqs/2, sum(Prcnt_seqs)/n_distinct(plant.species)/2 ),
         TAXON_EXPERT = plant.species) %>% 
  distinct(sample.id, plant.species, .keep_all = T)

@@

  mutate(TAXON_EXPERT = case_when(

    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 4 ~ plant.species, # 4 has a viola split

@@

    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', 
    sample.id == 19 ~ plant.species,
    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
    sample.id == 20 ~ plant.species,
    sample.id == 21 ~ plant.species,

    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 23 ~ plant.species, # has both Viola
    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,
    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,
    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,
    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 27 ~ plant.species, # has both Hydrophyllum
    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,
    sample.id == 29 ~ plant.species,
    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 30 ~ plant.species,
    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 32 ~ plant.species,
    sample.id == 33 ~ plant.species, # has both Viola

@@

    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 39 ~ plant.species, # has both hydrophyllum
    sample.id == 40 ~ plant.species,
    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
    sample.id == 41 ~ plant.species,
    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,
    sample.id == 44 ~ plant.species, # 44 has both delphinium

@@

    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 51 ~ plant.species, # 51 has a viola split
    sample.id == 52 ~ plant.species,
    sample.id == 54 ~ plant.species,

    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 55 ~ plant.species,

@@

   TRUE ~ as.character(TAXON_NEW)
  )) %>% 
  mutate(
    TAXON_EXPERT = str_replace_all(TAXON_EXPERT, '[.]', ' '),
    TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Geum ternatum', 'Geum triflorum'),
    TAXON_EXPERT = str_replace(
      TAXON_EXPERT, 'Arenaria globiflora|Minuartia recurva|Sagina procumbens', 'Eremogone congesta'),
    TAXON_EXPERT = if_else(doy < 171 & Genus == 'Cynoglossum', 
                           'Mertensia fusiformis', TAXON_EXPERT), 
    TAXON_EXPERT = if_else(doy < 190 & TAXON_EXPERT == 'Lupinus argenteus', 
                           'Lupinus sericeus', TAXON_EXPERT)) %>%

@@

rm(ranunc_splash, blst_expert)

# now we split up the reads assigned to conflicted groups  # e.g. Cynoglossum amplifolium in sample 65
b <- blst_reclass_expert %>% 
  filter(sample.id %in% dupe_samples) %>% 
  group_by(sample.id, Family) %>% 
  filter(n() >= 3) %>% 
  mutate(plant.species = na_if(plant.species, "Thalictrum.fendleri")) %>% 
  mutate(Seqs2distribute = sum(Prcnt_seqs_Reclass))

records2drop <- b %>% pull(RecordID)

b <- b %>% 
  drop_na(plant.species) %>% 
  mutate(Prcnt_seqs_Reclass = Seqs2distribute/n_distinct(TAXON_EXPERT)) %>% 
  select(-Seqs2distribute)

blst_reclass_expert <- blst_reclass_expert %>% 
  filter(!RecordID %in% records2drop) %>% 
  bind_rows(., b) %>% 
  arrange(sample.id, -Prcnt_seqs_Reclass)

rm(records2drop, b)

# 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? #
# seems i may not have either delphinium present in that week? and  if falls between two. 
# deal with sample 55 one off, split the delphinium gracile

s55 <- blst_reclass_expert %>% 
  filter(sample.id == 55) %>% 
  group_by(Genus) %>% 
  mutate(Seqs2distribute = sum(Prcnt_seqs_Reclass)) %>% 
  filter(TAXON_EXPERT != 'Delphinium gracile') %>% 
  mutate(Prcnt_seqs_Reclass = Seqs2distribute/n_distinct(TAXON_EXPERT)) %>% 
  select(-Seqs2distribute)

blst_reclass_expert <- blst_reclass_expert %>% 
  filter(sample.id != 55) %>% 
  bind_rows(., s55)  %>% 
  arrange(sample.id, -Prcnt_seqs_Reclass)

rm(s55)

# re class the totals to sum to 100%

blst_reclass_expert <- blst_reclass_expert %>% 
  group_by(sample.id) %>% 
  mutate(multiplier = 100/ sum(Prcnt_seqs_Reclass),
         Prcnt_seqs_Reclass2 = Prcnt_seqs_Reclass * multiplier)
